---
title: "RWorksheet_Buenvenida#1"
author: "Buenvenida"
date: "2024-09-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Set up a vector named age, consisting of 34, 28, 22, 36, 27, 18, 52, 39, 42, 29,
35, 31, 27, 22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 42, 53, 41,
51, 35, 24, 33, 41.

a. How many data points?


-There are 34 data points

b. Write the R code and its output.


```{r data points}
age <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27, 22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 42, 53, 41, 51, 35, 24, 33, 41) 
length(age)
```

2. Find the reciprocal of the values for age.
Write the R code and its output.

```{r }
age <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27, 22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 42, 53, 41, 51, 35, 24, 33, 41) 
reciprocals <- 1 / age

print(reciprocals)
```
3. Assign also new_age <- c(age, 0, age).
```{r}
age <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 
         35, 31, 27, 22, 37, 34, 19, 20, 57, 49, 
         50, 37, 46, 25, 17, 37, 42, 53, 41, 51, 
         35, 24, 33, 41) 

newage <- c(age, 0, age)

print(newage)

```
What happen to the new_age?

The `c()` function is used to combine vectors into one. In this case, it creates a new vector called `new_age`. This vector has 69 elements in total. It starts with the 34 elements from the original `age` vector, then adds a single `0`, and ends with the same 34 elements from the `age` vector again.

4. Sort the values for age.
Write the R code and its output.



```{r}
age <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 
         35, 31, 27, 22, 37, 34, 19, 20, 57, 49, 
         50, 37, 46, 25, 17, 37, 42, 53, 41, 51, 
         35, 24, 33, 41)
sortage <- sort(age)
print(sortage)
```

5. Find the minimum and maximum value for age.
```{r}
age <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 
         35, 31, 27, 22, 37, 34, 19, 20, 57, 49, 
         50, 37, 46, 25, 17, 37, 42, 53, 41, 51, 
         35, 24, 33, 41)
min_age <- min(age)
max_age <- max(age)
print(paste("Minimum value:", min_age))
print(paste("Maximum value:", max_age))

```

6. Set up a vector named data, consisting of 2.4, 2.8, 2.1, 2.5, 2.4, 2.2, 2.5,
2.3, 2.5, 2.3, 2.4, and 2.7.

a. How many data points?


-There are 12 data points

b. Write the R code and its output.


```{r}
data <- c(2.4, 2.8, 2.1, 2.5, 2.4, 2.2, 2.5, 2.3, 2.5, 2.3, 2.4, 2.7)

length_data <- length(data)

print(length_data)

```

7. Generates a new vector for data where you double every value of the data. | 


```{r}
data <- c(2.4, 2.8, 2.1, 2.5, 2.4, 2.2, 2.5, 2.3, 2.5, 2.3, 2.4, 2.7)
doubled <- data * 2
print(doubled)
```
What happen to the data?


- Each element is  multiplied by 2
\pagebreak
8. Generate a sequence for the following scenario:


8.1 Integers from 1 to 100.


```{r}

seq1to100 <- seq(from = 1, to = 100, by = 1)

print(seq1to100)

```
8.2 Numbers from 20 to 60


```{r}

seq20to60 <- seq(from = 20, to = 60, by = 1)

print(seq20to60)

```

8.3 Mean of numbers from 20 to 60

```{r}

numbers <- 20:60

mean <- mean(numbers)

print(mean)

```

8.4 Sum of numbers from 51 to 91

```{r}

numbers <- 51:91

mean <- mean(numbers)

print(mean)

```

8.5 Integers from 1 to 1,000
```{r}

seq1to1000 <- seq(from = 1, to = 1000, by = 1)

print(seq1to1000)

```
a. How many data points from 8.1 to 8.4?


-There are 143 data points

b. Write the R code and its output from 8.1 to 8.4.


8.1


```{r}
seq1to100 <- seq(from = 1, to = 100, by = 1)

print(seq1to100)

```


8.2


```{r}

seq20to60 <- seq(from = 20, to = 60, by = 1)

print(seq20to60)

```

8.3


```{r}
numbers <- 20:60

mean <- mean(numbers)

print(mean)


```

8.4

```{r}

numbers <- 51:91

mean <- mean(numbers)

print(mean)

```

\pagebreak
c. For 8.5 find only maximum data points until 10.


```{r}

seq1to1000 <- seq(from = 1, to = 1000, by = 1)

first10 <- seq1to1000[1:10]

max <- max(first10)

print(max)

```


9. *Print a vector with the integers between 1 and 100 that are not divisible by 3, 5 and
7 using filter option.

```{r}

sequence <- seq(1, 100)

filtered <- Filter(function(i) { all(i %% c(3, 5, 7) != 0) }, sequence)

print(filtered)

```


10. Generate a sequence backwards of the integers from 1 to 100.
Write the R code and its output.

```{r}
sequence <- seq(from = 1, to = 100, by = 1)

backwards <- rev(sequence)

print(backwards)

```


11. List all the natural numbers below 25 that are multiples of 3 or 5.
Find the sum of these multiples.

a. How many data points from 10 to 11?


-there are 112 data points

b. Write the R code and its output from 10 and 11.

```{r}
limit <- 25

numbers <- 1:(limit-1)

multiples <- numbers[numbers %% 3 == 0 | numbers %% 5 == 0]

summul <- sum(multiples)

print(multiples)        
print(summul)    

```


12. Statements can be grouped together using braces ‘{’ and ‘}’. A group of statements
is sometimes called a block. Single statements are evaluated when a new line is typed
at the end of the syntactically complete statement. Blocks are not evaluated until a
new line is entered after the closing brace.


Enter this statement:
x <- {0 + x + 5 +}

Describe the output:


Error: unexpected '}' in "x <- {0 + x + 5 +}"
there is a syntax issue with the statement provided. 


13. *Set up a vector named score, consisting of 72, 86, 92, 63, 88, 89, 91, 92, 75,
75 and 77. To access individual elements of an atomic vector, one generally uses the
x[i] construction.
Find x[2] and x[3]. Write the R code and its output.

```{r}
score <- c(72, 86, 92, 63, 88, 89, 91, 92, 75, 75, 77)

element2 <- score[2]
element3 <- score[3]


print(element2)
print(element3)

```


14. *Create a vector a = c(1,2,NA,4,NA,6,7).


a. Change the NA to 999 using the codes print(a,na.print="-999").
b. Write the R code and its output. Describe the output.

```{r}
a <- c(1, 2, NA, 4, NA, 6, 7)

print(a, na.print = "-999")

a[is.na(a)] <- 999

print(a)

```

The print(a, na.print = "-999") statement allows you to display NA values as -999 in the output, but does not change the vector.


To permanently replace NA with 999, the indexing method used a[is.na(a)] <- 999.


15. A special type of function calls can appear on the left hand side of the assignment
operator as in > class(x) <- "foo".
Follow the codes below:
name = readline(prompt="Input your name: ")
age = readline(prompt="Input your age: ")
