---
title: "RWorksheet_Buenvenida#4a.Rmd"
author: "me"
date: "2024-10-23"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. The table below shows the data about shoe size and height. Create a data frame.
a. Describe the data.

```{r}
HouseHoldData <- data.frame (
  ShoeSize = c(6.5, 9.0, 8.5, 8.5, 10.5, 7.0, 9.5, 9.0, 13.0, 7.5, 10.5, 8.5, 12.0, 10.5, 13.0, 11.5, 8.5, 5.0, 10.0, 6.5, 7.5, 8.5, 10.5, 8.5, 10.5,  11.0, 9.0, 13.0),
  
  Height = c(66.0, 68.0, 64.5, 65.0, 70.0, 64.0, 70.0, 71.0, 72.0, 64.0, 74.5, 67.0, 71.0, 71.0, 77.0, 72.0, 59.0, 62.0, 72.0, 66.0, 64.0, 67.0, 73.0, 69.0, 72.0, 70.0, 69.0, 70.0),
  
  Gender = c("F", "F", "F", "F","M", "F", "F", "F","M","F","M","F","M","M", "M", "M", "F", "F", "M", "F", "F","M", "M", "F","M","M", "M", "M")
)

HouseHoldData
```

b. Create a subset by males and females with their corresponding shoe size and height.
What its result? Show the R scripts.

```{r}
males <- subset(HouseHoldData, Gender == "M")
females <- subset(HouseHoldData, Gender == "F")
males
females
```

c. Find the mean of shoe size and height of the respondents. Write the R scripts and its
result.



```{r}
mean_shoe_size <- mean(HouseHoldData$ShoeSize)

mean_height <- mean(HouseHoldData$Height)

mean_shoe_size
mean_height
```


d. Is there a relationship between shoe size and height? Why?

there is a moderate positive correlation between shoe size and height, indicating that individuals with larger shoe sizes tend to be taller.


2. Construct character vector months to a factor with factor() and assign the result to
factor_months_vector. Print out factor_months_vector and assert that R prints out
the factor levels below the actual values.

```{r}
months_vector <- c("March", "April", "January", "November", "January",
                   "September", "October", "September", "November", "August",
                   "January", "November", "November", "February", "May", "August",
                   "July", "December", "August", "August", "September", "November", 
                   "February", "April")
factor_months_vector <- factor(months_vector)
months_vector

```


3. Then check the summary() of the months_vector and factor_months_vector. | Inter-
pret the results of both vectors. Are they both equally useful in this case?

```{r}
summary_months_vector <- summary(months_vector)
summary_factor_months_vector <- summary(factor_months_vector)

summary_months_vector
summary_factor_months_vector
```

4. Create a vector and factor for the table below.
```{r}
direction <- c("East", "West", "North")

frequency <- c(1, 4, 3)

factord <- direction
new_order <- factor(factord, levels = c("East", "West", "North"))

new_order

```

5. 
```{r}
#a
data <- read.table("import_march.csv", header = TRUE, sep = ",")
```
```{r}
#b
data
```

6.Full Search
```{r}
exhaustiveS <- function() {
  number <- suppressWarnings(as.integer(readline(prompt = "Select a number between 1 and 50: ")))
  if (is.na(number)) {
    print("Invalid input. Please enter a number.")
  } else if (number < 1 || number > 50) {
    print("The number selected is beyond the range of 1 to 50")
  } else if (number == 20) {
    print(TRUE)
  } else {
    print(number)
  }
}
exhaustiveS()
```


7. Change

```{r}
min_bills <- function(price) {
  bills <- c(1000, 500, 200, 100, 50)
  count <- 0
  for (bill in bills) {
    
    while (price >= bill) {
      price <- price - bill  
      count <- count + 1   
    }
  }
  return(count)
}
snack_price <- 2700
cat("Minimum number of bills needed:", min_bills(snack_price), "\n")
```
8.

```{r}
# a
grades <- data.frame(
  Name = c("Annie", "Thea", "Steve", "Hanna"),
  Grade1 = c(85, 65, 75, 95),
  Grade2 = c(65, 75, 55, 75),
  Grade3 = c(85, 90, 80, 100),
  Grade4 = c(100, 90, 85, 90)
)
print(grades)
```
```{r}
# b
for (i in 1:nrow(grades)) {
  avg <- sum(grades[i, 2:5]) / 4
  print(paste(grades$Name[i], "'s average grade this semester is", avg))
}
```
```{r}
# c
for (j in 2:5) {
  avg_test <- mean(grades[,j])
  if (avg_test < 80) {
    print(paste("The", j-1, "test was difficult with an average score of", avg_test))
  }
}
```
```{r}
# d
for (i in 1:nrow(grades)) {
  highest_score <- grades[i, 2]  
  for (j in 3:5) {  
    if (grades[i, j] > highest_score) {
      highest_score <- grades[i, j]
    }
  }
  if (highest_score > 90) {
    print(paste(grades$Name[i], "'s highest grade this semester is", highest_score))
  }
}
```



