---
title: "RWorksheet_Buenvenida#4c.Rmd"
author: "me"
date: "2024-11-11"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Use the dataset mpg

a. Show your solutions on how to import a csv file into the environment.

```{r}
mpgdata <- read.csv("mpg.csv")
mpgdata
```

b. Which variables from mpg dataset are categorical?

manufacturer - The car manufacturer (e.g., Audi, Chevrolet).

 - model
 - trans
 - drv
 - cyl
 - fl
 - class

c. Which are continuous variables?

  - displ
 - year
 - cyt
 - hwy
 
2. Which manufacturer has the most models in this data set? Which model has the most variations?
Show your answer.

```{r}
library(dplyr)

manufacturer_model_count <- mpgdata %>%
  group_by(manufacturer) %>%
  summarise(distinct_models = n_distinct(model)) %>%
  arrange(desc(distinct_models))

top_manufacturer <- manufacturer_model_count[1, ]

model_variation_count <- mpgdata %>%
  group_by(model) %>%
  summarise(variation_count = n()) %>%
  arrange(desc(variation_count))

top_model <- model_variation_count[1, ]

top_manufacturer
top_model
```



a. Group the manufacturers and find the unique models. Show your codes and result.

```{r}

library(dplyr)

manufacturer_unique_models <- mpgdata %>%
  group_by(manufacturer) %>%
  summarise(models_list = list(unique(model))) %>%
  arrange(manufacturer)

print(manufacturer_unique_models)

```

b. Graph the result by using plot() and ggplot(). Write the codes and its result.


```{r}
library(dplyr)

manufacturer_model_counts <- mpgdata %>%
  group_by(manufacturer) %>%
  summarise(model_count = n_distinct(model)) %>%
  arrange(desc(model_count))

barplot(manufacturer_model_counts$model_count, 
        names.arg = manufacturer_model_counts$manufacturer, 
        las = 2, 
        col = "lightblue", 
        main = "Number of Unique Models by Manufacturer",
        xlab = "Manufacturer", 
        ylab = "Number of Unique Models")

library(ggplot2)
ggplot(manufacturer_model_counts, aes(x = reorder(manufacturer, model_count), y = model_count)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  coord_flip() +
  labs(title = "Number of Unique Models by Manufacturer", 
       x = "Manufacturer", 
       y = "Number of Unique Models") +
  theme_minimal()
```


2. Same dataset will be used. You are going to show the relationship of the modeland the manufacturer.

a. What does ggplot(mpg, aes(model, manufacturer)) + geom_point() show?
The `ggplot(mpg, aes(model, manufacturer)) + geom_point()` creates a scatter plot showing the distribution of car models across manufacturers, which can help identify patterns in model representation for each manufacturer.

b. For you, is it useful? If not, how could you modify the data to make it more informative?

It is useful; however, modifying the data to show the count of unique models per manufacturer in a bar plot would provide clearer insights into the number of models per manufacturer.

3. Plot the model and the year using ggplot(). Use only the top 20 observations. Write the codes and
its results.

```{r}
library(ggplot2)
library(dplyr)

top_mpg_by_model <- mpgdata %>%
  arrange(desc(cty)) %>%
  head(20)

ggplot(top_mpg_by_model, aes(x = model, y = year)) +
  geom_point() +
  labs(title = "Top 20 Models by Year", x = "Model", y = "Year") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

 4. Using the pipe (%>%), group the model and get the number of cars per model. Show codes and its result
 
```{r}
library(dplyr)

model_car_counts <- mpg %>%
  group_by(model) %>%
  summarise(car_count = n()) %>%
  arrange(desc(car_count))

print(model_car_counts)

```

a. Plot using geom_bar() using the top 20 observations only. The graphs shoudl have a title, labels and
colors. Show code and results.

```{r}
library(dplyr)
library(ggplot2)

top_car_models <- mpgdata %>%
  group_by(model) %>%
  summarise(car_count = n()) %>%
  arrange(desc(car_count)) %>%
  slice_head(n = 20)

ggplot(top_car_models, aes(x = reorder(model, car_count), y = car_count, fill = model)) +
  geom_bar(stat = "identity") +
  coord_flip() +  
  labs(title = "Top 20 Car Models by Number of Cars",
       x = "Car Model",
       y = "Number of Cars",
       fill = "Model") +
  theme_minimal() +
  scale_fill_viridis_d() 

```

b. Plot using the geom_bar() + coord_flip() just like what is shown below. Show codes and its result.
```{r}
library(viridis)
library(viridisLite)
top_car_model_counts <- mpgdata %>%
  group_by(model) %>%
  summarise(car_count = n()) %>%
  arrange(desc(car_count)) %>%
  slice_head(n = 20)

ggplot(top_car_model_counts, aes(x = reorder(model, car_count), y = car_count, fill = model)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Top 20 Car Models by Number of Cars",
       x = "Car Model",
       y = "Number of Cars") +
  theme_minimal() +
  scale_fill_viridis_d()

```

5. Plot the relationship between cyl - number of cylinders and displ - engine displacement using
geom_point with aesthetic color = engine displacement. Title should be “Relationship between
No. of Cylinders and Engine Displacement”.

```{r}
ggplot(mpgdata, aes(x = cyl, y = displ, color = displ)) +
  geom_point(size = 3, alpha = 0.7) +  
  labs(title = "Relationship between No. of Cylinders and Engine Displacement",
       x = "Number of Cylinders",
       y = "Engine Displacement") +
  theme_minimal() +
  scale_color_viridis_c() 
```

a. How would you describe its relationship? Show the codes and its result.

The plot shows a positive correlation between the number of cylinders (`cyl`) and engine displacement (`displ`), with vehicles having more cylinders generally having larger engine displacements. The color gradient highlights this trend.

6. Import the traffic.csv onto your R environment.

```{r}
library(readr)
traffic_records <- read_csv("traffic.csv")
head(traffic_records)

```

a. How many numbers of observation does it have? What are the variables of the traffic dataset the Show
your answer.

num_traffic_records <- nrow(traffic_records)
traffic_variables <- colnames(traffic_records)
num_traffic_records
traffic_variables


b. subset the traffic dataset into junctions. What is the R codes and its output?

```{r}
junction_traffic_records <- subset(traffic_records, Junction == TRUE)
head(junction_traffic_records)

```

c. Plot each junction in a using geom_line(). Show your solution and output.

library(ggplot2)

ggplot(traffic_records, aes(x = DateTime, y = Vehicles, color = factor(Junction))) +
  geom_line() +
  labs(title = "Vehicle Counts at Junctions Over Time",
       x = "Date and Time",
       y = "Number of Vehicles",
       color = "Junction") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


7. From alexa_file.xlsx, import it to your environment
a. How many observations does alexa_file has? What about the number of columns? Show your solution
and answer.

```{r}
library(readxl)

alexa_traffic_data <- read_excel("alexa_file.xlsx")
data_dimensions <- dim(alexa_traffic_data)
num_rows_data <- data_dimensions[1]
num_columns_data <- data_dimensions[2]

num_rows_data
num_columns_data

```


b. group the variations and get the total of each variations. Use dplyr package. Show solution and
answer.

```{r}
library(dplyr)

variation_summary <- alexa_traffic_data %>%
  group_by(variation) %>%
  summarise(total_count = n())

variation_summary

```


c. Plot the variations using the ggplot() function. What did you observe? Complete the details of the
graph. Show solution and answer.

```{r}
library(ggplot2)

ggplot(variation_summary, aes(x = variation, y = total_count, fill = variation)) +
  geom_bar(stat = "identity") +
  labs(title = "Total Counts for Each Variation", x = "Variation", y = "Total Count") +
  theme_minimal()

```


d. Plot a geom_line() with the date and the number of verified reviews. Complete the details of the
graphs. Show your answer and solution.

```{r}
alexa_traffic_data$verified_reviews <- as.numeric(alexa_traffic_data$verified_reviews)

reviews_by_date_summary <- alexa_traffic_data %>%
  group_by(date) %>%
  summarise(total_verified_reviews = sum(verified_reviews, na.rm = TRUE))

library(ggplot2)

ggplot(reviews_by_date_summary, aes(x = date, y = total_verified_reviews)) +
  geom_line(color = "lightblue") +
  labs(title = "Verified Reviews Over Time", x = "Date", y = "Number of Verified Reviews") +
  theme_minimal()

```

e. Get the relationship of variations and ratings. Which variations got the most highest in rating? Plot a
graph to show its relationship. Show your solution and answer.
```{r}
variation_ratings_summary <- alexa_traffic_data %>%
  group_by(variation) %>%
  summarise(average_rating = mean(rating, na.rm = TRUE))

ggplot(variation_ratings_summary, aes(x = variation, y = average_rating, fill = variation)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Rating by Variation", x = "Variation", y = "Average Rating") +
  theme_minimal()

```


